\section{Installation}

\subsection{Prerequisites}

\begin{itemize}
\item Python 1.5.2 or later.

\item C compiler

The \module{Sybase} package contains an extension module written in C.
The extension module is used by the Python wrapper module
\module{Sybase.py}.

\item Sybase client libraries

The \module{Sybase} package uses the Sybase client libraries which
should have come as part of your Sybase installation.

If you are using Linux, Sybase provide a free version of their
database which can be downloaded from \url{http://linux.sybase.com/}.

\item mxDateTime

If the \module{mxDateTime} package is installed the \module{Sybase}
module can return \code{datetime} values as \class{DateTime} objects.
If \module{mxDateTime} is not present the module will use the
\class{DateTime} object defined in the \module{sybasect} extension
module.
\end{itemize}
\subsection{Installing}

The \module{Sybase} package the \module{distutils} package so all you
need to do is type the following command as root:

\begin{verbatim}
python setup.py install
\end{verbatim}

If you have problems with this step, edit the \texttt{setup.py} file
to specify where Sybase is installed and the name of the client
libraries.  Make sure that you contact the package author so that the
installation process can be made more roibust for other people.

\subsection{Testing}

The most simple way to test the \module{Sybase} package is to run a
test application.  The arguments to the \function{Sybase.connect()}
function are \var{server} (from the interfaces file), \var{username},
\var{password}, and \var{database}.  The \var{database} argument is
optional.  Make sure you substitute values that will work in your
environment.

\begin{verbatim}
>>> import Sybase
>>> db = Sybase.connect('SYBASE', 'sa', '')
>>> c = db.cursor()
>>> c.callproc('sp_help')
>>> for t in c.description:
...     print t
... 
('Name', 0, 0, 30, 0, 0, 0)
('Owner', 0, 0, 30, 0, 0, 32)
('Object_type', 0, 0, 22, 0, 0, 32)
>>> for r in c.fetchall():
...     print r
... 
('spt_datatype_info', 'dbo', 'user table')
('spt_datatype_info_ext', 'dbo', 'user table')
('spt_limit_types', 'dbo', 'user table')
  :
  :
('sp_prtsybsysmsgs', 'dbo', 'stored procedure')
('sp_validlang', 'dbo', 'stored procedure')
>>> c.nextset()
1
>>> for t in c.description:
...     print t
... 
('User_type', 0, 0, 15, 0, 0, 0)
('Storage_type', 0, 0, 15, 0, 0, 0)
('Length', 6, 0, 1, 0, 0, 0)
('Nulls', 11, 0, 1, 0, 0, 0)
('Default_name', 0, 0, 15, 0, 0, 32)
('Rule_name', 0, 0, 15, 0, 0, 32)
\end{verbatim}
