\chapter{Low Level Sybase Access}

\localmoduletable

\section{\module{sybasect} --- Interface to Sybase-CT library}

\declaremodule[sybasect]{extension}{sybasect}

\modulesynopsis{Interface to Sybase-CT library.}

This is not a complete reference to the Sybase CT library.  Sybase
produce excellent documentation which fully describes the use of the
CT library.

This document describes how to access the Sybase CT library while
using this module.

The \module{sybasect} extension contains the following:

\subsection{Types}

\begin{datadesc}{ContextType}
The type of \code{CS_CONTEXT} objects which wrap the Sybase
\code{CS_CONTEXT} structure pointer.
\end{datadesc}

\begin{datadesc}{ConnectionType}
The type of \code{CS_CONNECTION} objects which wrap the Sybase
\code{CS_CONNECTION} structure pointer.
\end{datadesc}

\begin{datadesc}{CommandType}
The type of \code{CS_COMMAND} objects which wrap the Sybase
\code{CS_COMMAND} structure pointer.
\end{datadesc}

\begin{datadesc}{BlkDescType}
The type of \code{CS_BLKDESC} objects which wrap the Sybase
\code{CS_BLKDESC} structure pointer.
\end{datadesc}

\begin{datadesc}{DataFmtType}
The type of \code{CS_DATAFMT} objects which wrap the Sybase
\code{CS_DATAFMT} structure.
\end{datadesc}

\begin{datadesc}{IODescType}
The type of \code{CS_IODESC} objects which wrap the Sybase
\code{CS_IODESC} structure.
\end{datadesc}

\begin{datadesc}{ClientMsgType}
The type of \code{CS_CLIENTMSG} objects which wrap the Sybase
\code{CS_CLIENTMSG} structure.
\end{datadesc}

\begin{datadesc}{ServerMsgType}
The type of \code{CS_SERVERMSG} objects which wrap the Sybase
\code{CS_SERVERMSG} structure.
\end{datadesc}

\begin{datadesc}{DataBufType}
The type of data buffers for sending and receiving data to and from
Sybase.  The type of object returned by \code{DataBuf('hello')}.
\end{datadesc}

\begin{datadesc}{NumericType}
The type used to store Sybase \code{CS_NUMERIC} and \code{CS_DECIMAL}
data values.
\end{datadesc}

\subsection{Functions}

\begin{funcdesc}{cs_ctx_alloc}{\optional{version \code{= CS_VERSION_100}}}
Execute the Sybase-CT function \function{cs_ctx_alloc()}.  The result
is a tuple containing the Sybase result code and a new instance of the
\class{CS_CONTEXT} class.  \code{None} is returned as the
\class{CS_CONTEXT} object when the result code is not
\code{CS_SUCCEED}.

\begin{verbatim}
from sybasect import *
status, ctx = cs_ctx_alloc(CS_VERSION_100)
if status != CS_SUCCEED:
    raise 'cs_ctx_alloc'
\end{verbatim}
\end{funcdesc}

\begin{funcdesc}{cs_ctx_global}{\optional{version \code{= CS_VERSION_100}}}
Execute the Sybase-CT function \function{cs_ctx_global()}.  The result
is a tuple containing the Sybase result code and a new instance of the
\class{CS_CONTEXT} class.  \code{None} is returned as the
\class{CS_CONTEXT} object when the result code is not
\code{CS_SUCCEED}.

\begin{verbatim}
from sybasect import *
status, ctx = cs_ctx_global(CS_VERSION_100)
if status != CS_SUCCEED:
    raise 'cs_ctx_global'
\end{verbatim}
\end{funcdesc}

\begin{funcdesc}{DataBuf}{obj}
Return a new instance of the \class{DataBuf} class.  The \var{obj}
argument is used to initialise the \class{DataBuf} object.

For all types of \var{obj} other than \class{CS_DATAFMT} a buffer will
be initialised which contains a single value.

\begin{verbatim}
buf = DataBuf('hello')
if cmd.ct_param(buf) != CS_SUCCEED:
    raise 'ct_param'
\end{verbatim}

When \var{obj} is a \class{CS_DATAFMT} object an empty buffer will be
created according to the attributes of the \class{CS_DATAFMT} object.
It is most common to create and bind a buffer in a single operation
via the \method{ct_bind()} method of the \class{CS_COMMAND} class.

For example, the following code creates a set of buffers for
retrieving 16 rows at a time.  Note that it is your responsibility to
ensure that the buffers are not released until they are no longer
required.

\begin{verbatim}
status, num_cols = cmd.ct_res_info(CS_NUMDATA)
if status != CS_SUCCEED:
    raise 'ct_res_info'
bufs = []
for i in range(num_cols):
    status, fmt = cmd.ct_describe(i + 1)
    if status != CS_SUCCEED:
        raise 'ct_describe'
    fmt.count = 16
    status, buf = cmd.ct_bind(i + 1, fmt)
    if status != CS_SUCCEED:
        raise 'ct_bind'
    bufs.append(buf)
\end{verbatim}
\end{funcdesc}

\begin{funcdesc}{numeric}{obj \optional{, precision \code{= -1}} \optional{, scale \code{= -1}}}
Return a new instance of the \class{NumericType} class.

The \var{obj} argument can accept any of the following types;
\code{IntType}, \code{LongType}, \code{FloatType}, \code{StringType},
or \code{NumericType}.

If greater than or equal to zero the \var{precision} and \var{scale}
arguments are used as the \member{precision} and \member{scale}
attributes of the \code{CS_DATAFMT} which is used in the Sybase
\function{cs_convert()} function to create the new \class{NumericType}
object.  The default values for these arguments depends upon the type
being converted to \class{NumericType}.

\begin{longtable}{l|l|l}
Type & \var{precision} & \var{scale} \\
\hline
\code{IntType}     & \code{16} & \code{0} \\
\code{LongType}    & \# of digits in \code{str()} conversion & \code{0} \\
\code{FloatType}   & \code{CS_MAX_PREC} & \code{12} \\
\code{StringType}  & \# digits before decimal point & \# digits after decimal point \\
\code{NumericType} & input precision & input scale \\
\end{longtable}
\end{funcdesc}

\begin{funcdesc}{sizeof_type}{type_code}
Returns the size of the type identified by the Sybase type code
specified in the \var{type_code} argument.

The values expected for the \var{type_code} argument things like;
\code{CS_CHAR_TYPE}, \code{CS_TINYINT_TYPE}, etc.
\end{funcdesc}

\begin{funcdesc}{CS_DATAFMT}{}
Return a new instance of the \class{DataFmtType} class.  This is used
to wrap the Sybase \code{CS_DATAFMT} structure.
\end{funcdesc}

\begin{funcdesc}{CS_IODESC}{}
Return a new instance of the \class{IODescType} class.  This is used
to wrap the Sybase \code{CS_IODESC} structure.
\end{funcdesc}

\begin{funcdesc}{CS_LAYER}{msgnumber}
Return the result of applying the Sybase \code{CS_LAYER} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_ORIGIN}{msgnumber}
Return the result of applying the Sybase \code{CS_ORIGIN} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_SEVERITY}{msgnumber}
Return the result of applying the Sybase \code{CS_SEVERITY} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\begin{funcdesc}{CS_NUMBER}{msgnumber}
Return the result of applying the Sybase \code{CS_NUMBER} macro to the
\var{msgnumber} argument.
\end{funcdesc}

\input{ctxobj}
\input{connobj}
\input{cmdobj}
\input{msgsobj}
\input{datafmtobj}
\input{databufobj}
\input{iodescobj}
\input{blkdescobj}
